---
title: "R-Ladies Workshop - Data Transformation, Wrangling and Communication"
author: "Divya Seernani"
date: "3rd July 2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
always_allow_html: yes
---

# Revising some Tidy Concepts
Although we do need to stick to either, Base R or Tidyverse, let us start by revising some concepts from the tidyverse universe.
Here is a cheatsheet - <https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf>

### Some old concepts - Data Transformation
Recall the data transformation workshop. We will once again need dplyr.

### Some new concepts - Data Wrangling
Open the cheatsheet. Let's go through when we can use gather, spread, separate and unite from the tidyr package.

### Data Visualisation
ggplot, BUT we also learnt about plotly last time, so let's try this again. It works best with html files such as the one we are trying to create here.

### Teaching and Communication
We will create this document, everything from this point on, in R-Markdown

#The task for today is to explore which cadies are the most and least popular

Today we take data from a 2017 [survey](http://www.scq.ubc.ca/wp-content/uploads/2017/10/candyhierarchysurvey2017.pdf) about people's preferences for halloween candy.

This is so much candy data, seriously! You can find it here - <http://www.scq.ubc.ca/so-much-candy-data-seriously/>

Let's break these steps down into a combination of R code, its output, and some text explaining what is happening

First, open a new markdown file. Load the packages we will be needing today - knitr, tidyverse, ggplot and plotly. 

```{r libraries, include=FALSE, results='hide'}
library(tidyverse)
library(ggplot2)
library(plotly)

library(knitr)
opts_knit$set(global.par = TRUE)

```


###Data transformation and wrangling

The example below is one way of using data wrangling and data transformation tools to first create a tidy dataset and then explore what people's cady preferences are.

```{r data wrangling and transformation}

##### Adding R Code - use {r eval=FALSE} if you do not want the code to run

candyhierarchy2017 <- read.csv("E:/R-Ladies/R-Ladies/candyhierarchy2017.csv", row.names=NULL)

candyhierarchy2017_2<-select(candyhierarchy2017, 7:109)

candyhierarchy2017_Tidy<-gather(candyhierarchy2017_2, "Option", "Answer")

candyhierarchy2017_Tidy$Joy<-ifelse(candyhierarchy2017_Tidy$Answer=="JOY", 1, 0)
candyhierarchy2017_Tidy$Despair<-ifelse(candyhierarchy2017_Tidy$Answer=="DESPAIR", 1, 0)

candyhierarchy2017_Joy<-aggregate(Joy ~ Option, candyhierarchy2017_Tidy, sum)
candyhierarchy2017_Despair<-aggregate(Despair ~ Option, candyhierarchy2017_Tidy, sum)

candyhierarchy2017_Tidy2<-merge(candyhierarchy2017_Joy, candyhierarchy2017_Despair, by="Option")


```

### Back to ggplot2 and plotly to make an interactive plot exploring our candy data.

Markdown let's you embed plots:

```{r ggplot, echo=TRUE, fig.width=15, fig.height=7}

##Note that the `echo = FALSE` parameter prevents printing of the R code that generated the plot.

CandyPlot<-ggplot(candyhierarchy2017_Tidy2, aes(x=Joy, y=Despair)) + 
  geom_point(aes(col=Option)) + 
  labs(subtitle="Halloween Candy 2017", 
       y="Despair", 
       x="Joy", 
       title="Scatterplot", 
       caption = "From Candy Heirarchy Data")

ggplotly(CandyPlot)

```

### Finally, Knit the markdown file into an html document. HTML is needed here to present data from plotly. You could also knit to .doc or .pdf.

