---
title: "Exploratory data analysis"
author: "R-Ladies Freiburg // Julia Müller"
date: "2023-05-02"
output: html_document
---

Exploratory data analysis (EDA)
- explore data to see main trends
- visualise important characteristics and relationships
- more of a mindset than a fixed set of steps
- uses descriptive statistics and graphs

Data
- from Tidy Tuesday project
- demographic representation of artists through editions of popular art history textbooks in the US

```{r}
library(tidyverse)
library(viridis)

artists <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-17/artists.csv') %>% 
  drop_na() %>% 
  filter(artist_race != "N/A")

head(artists)
```

Convert to factors when applicable, and summarise the two exhibition columns:
```{r}
artists <- artists %>% 
  mutate(across(c(artist_nationality, artist_gender, artist_race, artist_ethnicity, book, artist_unique_id, artist_race_nwi), as_factor),
         total_exhibitions = moma_count_to_year + whitney_count_to_year) 

summary(artists)
```


# Descriptive statistics

## Summaries

Median space ratio per page:
```{r}
artists %>% 
  summarise(median(space_ratio_per_page_total))
```

median = "halfway" point of the data: 50% lies above this point, 50% lies below


Average space ratio per page:
```{r}
artists %>% 
  summarise(mean(space_ratio_per_page_total))
```

What's the difference between the mean and the median?
- Extreme values (very high/low values) can influence the mean, but they don't influence the median
- On the other hand: the mean incorporates information from *all* data points

Standard deviation:
```{r}
artists %>% 
  summarise(sd(space_ratio_per_page_total))
```
The standard deviation expresses the "spread" of the data, i.e. how closely the data points are to the mean


All three:
```{r}
artists %>% 
  summarise(median(space_ratio_per_page_total), mean(space_ratio_per_page_total), sd(space_ratio_per_page_total))
```

With custom names:
```{r}
artists %>% 
  summarise("median space ratio" = median(space_ratio_per_page_total), 
            "mean space ratio" = mean(space_ratio_per_page_total), 
            "standard deviation" = sd(space_ratio_per_page_total))
```

Showing the minimum and maximum values (or range):
```{r}
artists %>% 
  summarise(min(space_ratio_per_page_total),
            max(space_ratio_per_page_total))
```


### Try it
What's higher - the average number of exhibitions in MOMA or Whitney? Also take a look at the relevant standard deviations.
```{r}

```


## Grouped summaries

We can calculate descriptive statistics separately for each level of a category:
```{r}
artists %>% 
  group_by(artist_gender) %>% 
  summarise(mean(space_ratio_per_page_total),
            sd(space_ratio_per_page_total))
```

```{r}
artists %>% 
  group_by(artist_gender) %>% 
  summarise(max(space_ratio_per_page_total))
```

Which person got the most pages?
```{r}
artists %>% 
  filter(space_ratio_per_page_total == max(space_ratio_per_page_total))
```

Which woman got the most coverage?
```{r}
artists %>% 
  filter(artist_gender == "Female") %>% 
  filter(space_ratio_per_page_total == max(space_ratio_per_page_total))
```

Number of exhibitions by nationality:
```{r}
artists %>% 
  distinct(artist_unique_id, .keep_all = TRUE) %>% 
  group_by(artist_nationality) %>% 
  summarise("number of exhibitions" = sum(total_exhibitions)) %>% 
  arrange(desc(`number of exhibitions`))
```


## Counts 
How many artists of which races are included in the textbooks?
```{r}
artists %>% 
  distinct(artist_unique_id, .keep_all = TRUE) %>% # to only keep unique artists, but also retain the information in all other columns
  count(artist_race)
```


How could you see the average pages and average exhibitions by the artist's nationality? We can combine this with `arrange()` to sort by most/least covered or exhibited nationalities.
```{r}
artists %>% 
  group_by(artist_nationality) %>% 
  summarise("average space ratio" = mean(space_ratio_per_page_total), 
            "average MOMA" = mean(moma_count_to_year), 
            "average Whitney" = mean(whitney_count_to_year)) %>% 
  arrange(desc(`average MOMA`)) # the quotation marks are necessary because the column name has a space
```


### Try it
Repeat the analysis - but with the artist's race instead of nationality. Then try grouping by both race and gender.
```{r}

```

```{r}
artists %>% 
  group_by(artist_race, artist_gender) %>% 
  summarise("average space ratio" = mean(space_ratio_per_page_total)) %>% 
  arrange(desc(`average space ratio`))
```

Also count how many artists of which race and gender are written about in the textbooks.
```{r}
artists %>% 
  distinct(artist_unique_id, .keep_all = TRUE) %>% 
  count(artist_race, artist_gender) %>% 
  arrange(desc(n))
```


# Graphs

## Space ratio by race and gender
```{r}
ggplot(artists) +
  aes(x = reorder(artist_race, -space_ratio_per_page_total), y = space_ratio_per_page_total, colour = artist_gender) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "artist's race",
       y = "space ratio per page",
       colour = "artist's gender",
       title = "Space given to artists in textbooks (1926 - 2020)") +
  scale_color_viridis(discrete = TRUE, option = "C")
```

## Space ratio by year and race
```{r}
ggplot(artists) +
  aes(x = year, y = space_ratio_per_page_total, colour = artist_race) +
  geom_jitter() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(y = "space ratio per page",
       colour = "artist's race",
       title = "Development of space given to artists by race") +
  theme(legend.position = "top") +
  scale_colour_viridis(discrete = TRUE, option = "B")
```

## Number of mentions by race and gender
```{r}
ggplot(artists) +
  aes(y = artist_race, fill = artist_gender) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(title = "Mentions in art textbooks by the artist's race and gender",
       fill = "") +
  scale_fill_viridis(discrete = TRUE, option = "G")
```

## Distinct artists by race and gender
```{r}
artists %>% 
  distinct(artist_unique_id, .keep_all = TRUE) %>% 
  ggplot() +
  aes(y = artist_race, fill = artist_gender) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(title = "Unique artists mentioned in textbooks by race and gender",
       fill = "") +
  scale_fill_viridis(discrete = TRUE, option = "G")
```

## Number of exhibitions
```{r}
artists %>% 
  group_by(artist_race, artist_gender) %>% 
  add_count(wt = total_exhibitions, name = "num_exhibitions") %>% 
  ungroup() %>% 
  distinct(artist_race, artist_gender, num_exhibitions) %>% 
  ggplot() +
  aes(x = reorder(artist_race, -num_exhibitions), y = num_exhibitions, fill = artist_gender) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(x = "artist's race",
       y = "exhibitions (MOMA and Whitney combined)",
       fill = "",
       title = "Number of exhibitions by artist's gender and race") +
  theme(legend.position = "top")
```

