---
title: "Best of 2021 - Data visualisation"
author: "Julia Müller"
date: "10 2 2022"
output: html_document
---

# Library and data

Source and more info:
https://github.com/rfordatascience/tidytuesday/blob/84c6abbadd5a038b99ca01da5aa377a07c8b46a7/data/2022/2022-01-18/readme.md
```{r}
library(tidyverse)
chocolate <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')

str(chocolate)
head(chocolate)

chocolate %>% 
  slice_sample(n = 10)
```

## Wrangling
```{r}
chocolate <- chocolate %>% 
  separate(col = "ingredients",
           into = c("number_ingredients", "list_ingredients"),
           sep = "- *") %>% 
  mutate(number_ingredients = as.numeric(number_ingredients),
         vanilla = case_when(
           str_detect(list_ingredients, "V") ~ "contains vanilla",
           TRUE ~ "no vanilla"),
         vanilla = as_factor(vanilla),
         cocoa_percent = str_remove(cocoa_percent, "%"),
         cocoa_percent = as.numeric(cocoa_percent),
         number_ingredients = case_when(
      is.na(number_ingredients) & cocoa_percent == 100 ~ 1,
      TRUE ~ number_ingredients
    ),
         across(where(is.character), as_factor)
    )

str(chocolate)
```


# Basic graphs

## Scatter plot: Cocoa percentage by chocolate rating
```{r}
(p1 <- chocolate %>% 
   drop_na(vanilla) %>% 
  ggplot() +
  aes(x = cocoa_percent, y = rating, colour = vanilla, fill = vanilla) +
  geom_jitter() +
  geom_smooth())
```

## Bar graph: Most common countries in data
```{r}
(p2 <- chocolate %>% 
  count(country_of_bean_origin) %>% 
  filter(country_of_bean_origin != "Blend" & n > 80) %>% 
  mutate(country_of_bean_origin = fct_reorder(country_of_bean_origin, n)) %>% 
  ggplot() +
  aes(x = n, y = country_of_bean_origin, fill = country_of_bean_origin) + 
  geom_col(show.legend = FALSE))
```


# Tips and tricks

## Lesser-known labels

Most users are familiar with the more common `labs()` arguments such as title, x, and y - but tags and captions are also very useful:
```{r}
(p1 <- p1 + 
  labs(title = "Do chocolates with a higher cocoa content taste better?",
       subtitle = "Do people like vanilla in their chocolate?",
       x = "% of cocoa in the chocolate bar",
       y = "rating (0 - 5)",
       tag = "Plot 1",
       caption = "Data source: Flavors of Cacao via TidyTuesday"))

(p2 <- p2 + 
  labs(title = "Where does chocolate come from?",
       subtitle = "Most frequent countries in the dataset",
       x = "frequency",
       y = "country",
       tag = "Plot 2",
       caption = "Data source: Flavors of Cacao via TidyTuesday"))
```

## Theme options
Workshop materials:
https://github.com/rladies/meetup-presentations_freiburg/blob/master/2021-10-13_Levelupggplot_theme/ggplot_theme_RLadiesFreiburg.Rmd

Video on Youtube:
https://www.youtube.com/watch?v=_DrfZzD_3nc&ab_channel=R-LadiesGlobal


Themes let you customize the "non-data components" of plots, such as its titles, labels, legends, etc. 
For p1, we'll center the title and subtitle, making the title bold and the subtitle italic and gray, decrease the tag's font size and put the legend at the top.
For p2, we'll delete the axis ticks and the colour of the panel background.

```{r}
p1 +
  theme_light() + # in-built presets
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, face = "italic", colour = "darkgray"),
    plot.tag = element_text(size = 10),
    legend.position = "top"
  )

p2 + 
  theme(
    axis.ticks = element_blank(),
    panel.background = element_rect(fill = "transparent")
  )
```


## Custom colours
Workshop materials:
https://github.com/rladies/meetup-presentations_freiburg/blob/master/2021-08-03_ggplot_colors/ggplot_colors_RLadiesFreiburg_KEY.Rmd

Video on Youtube:
https://www.youtube.com/watch?v=A4fP31h4roU&list=PLPwprT5wdzX7NVDl4oYQ7c2_6ox0_1fyr&index=6&t=9s&ab_channel=R-LadiesGlobal

R has several hundred inbuilt colour options:
```{r}
p1 +
  scale_color_manual(values = c("yellow", "green")) +
  scale_fill_manual(values = c("yellow", "green"))
```

...but also allows for custom colours with hex codes:
```{r}
p1 +
  scale_color_manual(values = c("yellow", "green")) +
  scale_fill_manual(values = c("yellow", "green"))
```

Or we can use palettes from other packages:
```{r}
p2 +
  viridis::scale_fill_viridis(discrete = TRUE)

p2 + 
  paletteer::scale_fill_paletteer_d("futurevisions::earth")
```


## Custom fonts
Workshop materials:
https://github.com/rladies/meetup-presentations_freiburg/blob/master/2021-11-10_ggplot_fonts/ggplot_fonts_RLadiesFreiburg.Rmd

Video on Youtube:
https://www.youtube.com/watch?v=t8sxun6UmWA&ab_channel=R-LadiesGlobal

As a quick example, we'll add a Google font (but note that it's also possible to use fonts that are installed on your computer):
```{r}
library(showtext)

font_add_google(name = "Indie Flower", # Name of the font as it is listed on Google fonts
                family = "indie") # Name we'll use in theme()

showtext_auto() # allows fonts to show up in plots
```

...then change the font in our plot:
```{r}
p2 +
  theme(text = element_text(family = "indie"))
```

Instead of having a separate legend, we can use colours in the plot title to explain what the two colours in p1 mean. The {ggtext} package also lets you use Markdown formatting for **bold** and *italic* words.
```{r}
library(ggtext)

(p1 <- p1 +
  theme_light() +
  scale_fill_manual(values = c("#2F4858", "#C34A36")) +
  scale_colour_manual(values = c("#2F4858", "#C34A36")) +
  labs(title = "Do chocolates with a **higher cocoa content** taste better?",
  subtitle = "Do these raters like
       <span style ='color:#C34A36'>vanilla in their chocolate</span>
       or
       <span style ='color:#2F4858'>do they prefer it without?</span>") +
   theme(plot.title = element_markdown(),
         plot.subtitle = element_markdown(),
        legend.position = "null"))
```


## Annotations
Workshop materials:
https://github.com/rladies/meetup-presentations_freiburg/blob/master/2021-12-08_ggplot_annotations/ggplot_annotations_KEY_RLadiesFreiburg.Rmd

Video on Youtube:
https://www.youtube.com/watch?v=If3LgW69Tx4&ab_channel=R-LadiesGlobal

Very high chocolate-content bars seem to not be everyone's cup of tea. We can point that out in a text box with `annotate(geom = "text")`:
```{r}
p1 +
  annotate(geom = "text",
           x = 95, y = 1.6,
           label = "Reviews for chocolate bars \nwith 100% cocoa \nare... mixed, at best") +
  geom_curve(
    aes(x = 95, y = 1.4, xend = 100, yend = 1.1),
    arrow = arrow(length = unit(0.1, "inch")), size = 0.6,
    color = "darkgray", curvature = 0.5
  )
```

Or we can annotate p1 to explain that chocolate bars with a cocoa content that is close to the average are rated the highest, using {ggforce} to draw a circle around these data points:
```{r}
library(ggforce)

p1 +
  geom_mark_circle(aes(
    fill = NA, colour = NA, # this lets us draw just one circle instead of one per vanilla/no vanilla data
    filter = cocoa_percent > 65 & cocoa_percent < 75 & rating > 2 & rating < 4, # filter to data that should be contained in the circle
    description = "There seems to be a sweet spot in terms of cocoa content: Chocolate bars with around 65-75% cocoa receive the best ratings."))
```


## Labels inside bars
Workshop materials (not recorded):
https://github.com/rladies/meetup-presentations_freiburg/blob/master/2021-08-24_GuidedTT_startrek/21_08_24_StarTrek.Rmd

Bar graphs can be labelled inside the bars rather than on the y-axis:
```{r}
p2 + 
  geom_text(aes(label = country_of_bean_origin,
                hjust = 1.1)) + # to ensure labels are inside the bars
  theme_void()
```


## Images instead of points
Workshop materials (not recorded):
https://github.com/rladies/meetup-presentations_freiburg/blob/master/2021-10-27_GuidedTT_pumpkins/21_10_27_Pumpkins.Rmd

Instead of dots to represent data points, `geom_image` in {ggimage} replaces them with a picture of your choice:
```{r}
library(ggimage)
img <- "choc.png"

chocolate %>% 
  filter(country_of_bean_origin == "Nicaragua") %>%
  ggplot() +
  aes(x = cocoa_percent, y = rating) +
  geom_image(aes(image = img),
             asp = 1.5) +
  labs(title = "Chocolate bars by rating, cocoa from Nicaragua only") +
  theme_light()
```


# Challenge ideas

- Try to recreate p1, but with salt content instead of vanilla. Use custom colours and any annotations you like!
- Pick a different image and recreate the last graph we made.
- Go back to our materials on the topic and figure out how to make the placement of labels (inside or outside of the bar) in p2 dependent on the data - i.e. place it inside for longer and outside for shorter bars.

