---
title: "Markdown Workshop"
author: "Divya and Elisa"
date: "23 5 2020"
output: 
  html_document:
    toc: True
    toc_float: true
    number_section: false
    theme: journal
    highlight: tango
params:
  tbi_age: !r readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv')
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Let's make a Markdown html together

You now have the rendered html text, let's discuss what went into making this text, i.e. what all should your text be able to do to render this document?

At the last Tidy Tuesday event, there was a discussion to make a kind of template for how we do Tidy Tuesday. In this document, I can explain to you how I went about creating my last Tidy Tuesday plot. I want to use the Tidy Tuesday data on Traumatic Brain Injury (TBI). I worked on this on our Meetup from 19.05.2020, although the data was posted earlier this year.

If you have a similar dataset, you can do so for that script. If not, just use mine :)

Something to help you along the way

- Shortcut to convert an r script to a markdown file
CTRL + Shift + k on windows OR Cmd + Shift + k on macOS

- Shortcut to create r code chunks
CTRL + ALT + i on windows OR Cmd + Option + I on macOS

## Starting off with R Markdown

You can start a new markdown file by going to file-> new file-> R Markdown. Give it a title and select HTML - you can always change these later. The document that shows up already has some information about Markdown. What does this tell us?

1. KnitR
2. Headers and Text
3. R-chunks
2. Echo function within chunks.

## Pre-processing

First I import my data. There were three datasets to choose from (or use together). I chose the one with age and number of cases of TBI in the year 2014.
I do not need to show you the output that says the csv has been included, so here, I can use echo=False, and include=FALSE. What is the difference?

```{r, echo=FALSE, include=FALSE}
# tbi_age <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv')
tbi_age <- params$tbi_age
```

The first thing I would do is simply look at the data. Open the table and look at it, use the str() and summary() functions. The question I ask here is - do I need to change anything to be able to plot my dataset?

```{r}
str(tbi_age)

```

You can already see the output here looks a bit untidy, so let's try and plot the summary in a table.

```{r}

summary<-summary(tbi_age)

knitr::kable(summary, caption = 'Summary Table of TBI_Age')

```

Because you can see my output, you can also see fr yourself that age is not a factor, and is honestly arranged a bit weirdly. If I want to plot age, I need age to be a factor. I would also like it to be ordered in an ascending order that makes sense while looking at it. So I make these two changes.

```{r}
tbi_age$age_group<-as.factor(tbi_age$age_group)
tbi_age$age_group <- ordered(tbi_age$age_group, levels = c("0-4", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75+"))

tbi_age<-na.omit(tbi_age)
```

## Making the Plot

I want to make a simple Bar graph, but for this Tidy Tuesday, I wanted to learn how to add a background image to my ggplot. For these, I need the following ggplot2, but also ggimage, magick and cowplot. I will also be using the Game of Thrones colour palette, so I'll call this package as well.

I do not need to show you how my packages run, so here, I can use _echo=False_, and _include=FALSE_.

```{r, echo=FALSE, include=FALSE}
library(ggplot2)
library(gameofthrones)
library(ggimage)
library(magick)
library(cowplot)
```

Now the plot itself took a bit of trial and error which I will not go into, but I will explain what I am doing here. 

1. I make a ggplot **bar graph** with a transperent background (in theme), using the House Lannister colour scheme and changing the text size in the axis so they are visible against the background.  
2. I **import the picture** I want to use as backgroud, the dimentions of which you see below. This is too loud and will overpower the graph.

```{r}
bar_plot<-ggplot(tbi_age, aes(x=age_group, y=number_est, fill=type)) + 
  geom_bar(position="dodge",stat="identity", width=1) + 
  scale_fill_got_d(option = "Lannister", direction = - 1)+
  labs(title="Tidy Tuesdays", 
       subtitle="Number of cases of Traumatic Brain Injury per Age-Group", 
       caption="R-Ladies Freiburg, 19.05.2020",
       y="Number of Cases of TBI", 
       x="Age Group")+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent",colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA),
    axis.text.x=element_text(angle=50, size=14, colour= "black", vjust=0.5),
    axis.text.y=element_text(size=14, colour= "black"),
    plot.title = element_text(size=20, face="bold", margin = margin(10, 0, 10, 0))
  )


img = image_read('presentation-figure/photo-1510698454686-1e2552e058e0.jpg')
#image_info(img)

```

3. I use **filters** from then 'magick' package to change its colours and blur it.  
1. I **remove scientific annotations** so the y-axis shows absolute values.  
1. Finally, I **add the background image** to the bar plot.

```{r}
img2<-image_negate(img)
img3<-image_blur(img2, 6, 5)

options(scipen=999)

ggdraw() +
  draw_image(img3) +
  draw_plot(bar_plot)  
```

So this was an easy ways for you to follow what I did and replicate my code.

# References  {.tabset}

## R-Ladies Oslo
[Reports to impress your boss - R-Ladies Oslo](https://www.youtube.com/watch?v=Xn5AmUf7gDQ)  

## R-Markdown definitive guide

[R-Markdown the definative guide](https://bookdown.org/yihui/rmarkdown/)


