---
title: "Adding images to ggplot with ggdraw and cowplot"
author: "Kyla McConnell"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot) #contains ggdraw
```

# Load data

- Tidy Tuesday animal crossing data -- customer reviews of 2020 released video game (fictional island with animal characters, from different species and personality types)
- character_reviews: all reviews that contain a character name, incl. character species and personality type

```{r, message=FALSE, warning=FALSE}
character_reviews <- read_csv("character_reviews.csv")
```

# Example 1: Add basic logo

Bar plot showing average grade by character. Color-coded (custom hex codes used here) to show personality types. Ordered in by descending average review score.

Make sure to save to a variable! 

```{r, message=FALSE, warning=FALSE}
(names_plot <- ggplot(data=character_grades, aes(x=reorder(name, -average_grade), y=average_grade, fill=personality)) +
  geom_col(show.legend=FALSE)+ 
  scale_fill_manual(values=c("#99d6ff", "#BFC2C6", "#C4E5C5", "#FFC9BF", "#95E3DF", "#F6F4CB", "#fac998", "#d6b2ab"))+
  xlab("Villager name") +
  ylab("Average review (out of 10)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family="serif"), panel.background=element_blank(), plot.title = element_text(family = "serif"), axis.text.y = element_text(family="serif"), axis.title=element_text(family="serif")))
```
Add image:

1. Find image URL (i.e. on Google)
2. Set up a ggdraw() object 
3. draw_plot(plot_name)
4. draw_image(url, x=, y=, scale=)
  - x and y co-ordinates from the axes, 0 is the middle, might need a bit of trial and error
  - scale to change the original size of the picture to make it fit the graph
  
```{r, message=FALSE, warning=FALSE}
ggdraw() +
  draw_plot(names_plot) +
  draw_image("https://vignette.wikia.nocookie.net/siivagunner/images/9/99/Animal_Crossing-_New_Horizons.jpg/revision/latest?cb=20190913005933", scale = .2) #Try x = 0.5, x = -0.5
```

```{r, message=FALSE, warning=FALSE}
ggdraw() +
  draw_plot(names_plot) +
  draw_image("https://vignette.wikia.nocookie.net/siivagunner/images/9/99/Animal_Crossing-_New_Horizons.jpg/revision/latest?cb=20190913005933", x=0.33, y=0.2, scale=.2)
```

# Example 2: Add headers to bar plots

Basic plot:

First, a little wrangling to produce a summary table with average ratings for each personality type (based on that character's name appearing in the review).

Bar chart showing the average for each personality type (again, custom hex codes for colors, a few changes to the font siyes, re-sizing axes and othher cosmetic changes)

Save to a variable!
```{r}
(species_plot <- character_reviews %>% 
  group_by(personality) %>% 
  summarize(average_grade = mean(average_grade)) %>%
  ggplot(aes(x=personality, y=average_grade, fill=personality))+
  geom_col(show.legend=FALSE)+ 
  xlab("Villager personality") +
  ylab("Average review (out of 10)")+
  ggtitle("How do villager's personalities affect user reviews in Animal Crossing?")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family="serif"), panel.background=element_blank(), plot.title = element_text(family = "serif", face="bold"), axis.text.y = element_text(family="serif"), axis.title=element_text(family="serif")) + 
  scale_fill_manual(values=c("#99d6ff", "#BFC2C6", "#C4E5C5", "#FFC9BF", "#95E3DF", "#F6F4CB", "#fac998", "#d6b2ab")) +
  scale_y_discrete(limits=c(0, 2, 4, 6, 8, 10))+
  ylim(0,10))
```
Add image toppers to each bar

1. Find image URLs (i.e. on Google) -> works best if they have transparent backgrounds or you remove the background first
2. Set up a ggdraw() object 
3. draw_plot(plot_name)
4. draw_image(url, x=, y=, scale=)

```{r}
ggdraw() +
  draw_plot(species_plot) +
  draw_image("https://villagerdb.com/images/villagers/thumb/robin.e227ada.png", scale=.5) #Try scaling down even further
```

  
```{r}
ggdraw() +
  draw_plot(species_plot) +
  draw_image("https://villagerdb.com/images/villagers/thumb/robin.e227ada.png", x = 0.31, y = 0.15, scale = .2) +
  draw_image("https://villagerdb.com/images/villagers/thumb/marshal.20eb375.png", x = 0.2, y = -0.05, scale = .2)+
  draw_image("https://villagerdb.com/images/villagers/thumb/flora.b45e15f.png", x = 0.09, y = 0.07, scale = .2)+
  draw_image("https://villagerdb.com/images/villagers/thumb/merengue.6609ed1.png", x = -0.03, y = 0.12, scale = .2) +
  draw_image("https://villagerdb.com/images/villagers/thumb/pashmina.8916368.png", x = 0.42, y = 0.08, scale = .2) +
  draw_image("https://villagerdb.com/images/villagers/thumb/beau.b905798.png", x = -0.12, y = 0.15, scale = .2) +
  draw_image("https://villagerdb.com/images/villagers/thumb/sterling.fef732a.png", x = -0.24, y = 0, scale = .2) +
  draw_image("https://villagerdb.com/images/villagers/thumb/del.5f28b25.png", x = -0.36, y = 0.1, scale = .2)
```
# Add a second plot as an inset

1. Save both plots as variables
2. Call ggdraw on the larger/main plot
3. draw_plot(other_plot, x=, y=, width=, height=)
  - where other_plot is your plot variable name
  - x and y are coordinnates, width/height for resizing
  
```{r}
#First I remove the old eader and y-axis label for clarity
species_plot <- species_plot + 
  ylab("")+
  ggtitle("")
  
ggdraw(names_plot) +
  draw_plot(species_plot, x = .52, y = .6, width = .4, height = .4)
```

And you can even add images to multiple plot layouts too!
```{r}
ggdraw(names_plot) +
   draw_plot(species_plot, x = .6, y = .55, width = .4, height = .4) +
  draw_image("https://vignette.wikia.nocookie.net/siivagunner/images/9/99/Animal_Crossing-_New_Horizons.jpg/revision/latest?cb=20190913005933", x= 0.02, y=0.3, scale=.2) +
  draw_image("https://villagerdb.com/images/villagers/medium/mitzi.7ed5f5f.png", x = 0.4, y = -0.18, scale = .25)
```

More info here:
https://wilkelab.org/cowplot/articles/drawing_with_on_plots.html
