---
title: "Guided Tidy Tuesday 06-05-21: Mario Kart"
author: "Kyla McConnell"
date: "7/6/2021"
output: html_document
---

# Guided Tidy Tuesday: Mario Kart 

```{r}
library(tidyverse)

records <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/records.csv')
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/drivers.csv')
```

## Look at data 
```{r}
glimpse(records)
summary(records)
str(records)
head(records)
```

```{r}
records %>% 
  mutate(track = as.factor(track),
         type = as.factor(type))

(records <- records %>% 
  mutate(across(where(is.character), as.factor)))
```


How did the world records develop over time?
Which track is the fastest?
For which track did the world record improve the most?
For how many tracks have shortcuts been discovered?
When were shortcuts discovered?
On which track does the shortcut save the most time?
Which is the longest standing world record?
Who is the player with the most world records?
Who are recent players?

Which loop is the most difficult?
Fastest tracks
Longest records 
Duration of record by track
Players vs number of shortcuts

```{r}
single_lap <- records %>% 
  filter(type == "Single Lap")

records %>% 
  group_by(track, type, shortcut) %>% 
  summarize(min_time = min(time)) %>% 
  ggplot() +
  aes(y = fct_reorder(track, desc(min_time)), x = min_time, fill = type) +
  geom_col(position = "dodge") +
  theme_minimal() + 
  labs(
    title = "Records on Mario Kart Tracks",
    y = "Record time", 
    x = "Track"
  ) +
  facet_wrap(type ~ shortcut)
```
Shortcuts:
How many have shortcuts?
When were they discovered?
How much time do they save?
```{r}
records %>% 
  filter(shortcut == "Yes") %>% 
  distinct(track)
```

```{r}
records %>% 
  mutate(year = lubridate::year(date)) %>% 
  filter(shortcut == "Yes") %>% 
  group_by(track) %>% 
  summarize(discovered = min(date)) %>% 
  arrange(discovered)
```


```{r}
records %>% 
  filter(shortcut == "Yes") %>% 
  count(player) %>% 
  arrange(desc(n))
```

```{r}
records %>% 
  group_by(track, shortcut) %>% 
  summarize(fastest = min(time)) %>% 
  pivot_wider(names_from = shortcut, values_from = fastest) %>% 
  mutate(Difference = No - Yes)
```

```{r}
kart <- drivers %>% 
  distinct(player, nation) %>% 
  full_join(records)

kart %>% 
  drop_na(nation) %>% 
ggplot(aes(y = fct_infreq(nation), fill = nation)) +
  geom_bar() +
  theme_minimal() +
    theme(
    legend.position = "None"
  ) +
  labs(
    title = "Records by country",
    y = "Country", 
    x = "# of Records"
  )
```

```{r}
records %>% 
  group_by(track, shortcut) %>% 
  summarise(fastest = min(time)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = shortcut, values_from = fastest) %>% 
  drop_na(Yes) %>% 
  mutate(time_saved = No - Yes,
         track = fct_reorder(track, -Yes)) %>% 
  ggplot() +
  aes(x = track) +
  geom_segment(aes(xend = track, 
                   y = Yes, yend = No), 
               color="grey") +
  geom_point(aes(y = Yes), 
             size = 3, colour = "darkblue") +
  geom_point(aes(y = No), 
             size = 3, colour = "darkgreen") +
  coord_flip() +
  theme_minimal()
```
