---
title: "Shiny: Hands-On"
author: "Kyla McConnell"
date: '2022-05-16'
output: html_document
---

# Shiny in a Nutshell

- Start Shiny from a template via File -> New Project -> New Directory -> Shiny Web Application
- Don't forget to load packages!
- You can preview Shiny in a new window or the Viewer pane but remember, this will keep R busy! 

## Three key parts of every Shiny app

### (1) ui
```{r}
ui <- fluidPage(
    selectInput(
        inputId = "country", 
        label = "Which country would you like to see?", 
        choices = c("Germany", "France", "Spain", "Italy", "Poland")
                ),
    plotOutput("medals")
)
```

- "User interface"

(A) Collects input from users:
  - drop-down list of options with *selectInput()*
  ```
    selectInput(inputId = "greeting", 
                label = "Who do you want to greet?", 
                choices = c("world" = "Hello, world!", "friends" = "Hi, friends!", "people" = "Hey, people!")
                )
  ```
  - box for typing input with *textInput()*
  ```
    textInput(inputId = "greeting", 
              label = "Who do you want to greet?", 
              value = "Type your name"
              )
  ```
  - numeric slider with *sliderInput()*
  ```
  sliderInput(inputID = "age", 
            label = "How old are you?", 
            min = 1, max = 80, value = 6
            )
  ```
 
(B) Shows output to users once it has been handled by the server

- The output object must first be generated (and named) by the server

  - lines of text with *textOutput()*
  ```
  textOutput("full_greeting")
  ```
  
  - summary table / tibble with *tableOutput()*
  ```
  tableOutput("medals")
  ```
  
  - ggplot / other plot with *plotOutput()* 
  ```
  plotOutput("medals")
  ```

(C) Both collecting input and showing output takes place within a *fluidPage()*, and note the commas! 


### (2) server
```{r}
server <- function(input, output, session) {
    output$full_greeting <- renderText(input$greeting)
}
```

- Takes the users input, does something to it, and returns it as an output object
- Input is read-only from the server
- Input has to be read into a reactive context like `renderText()`, `renderPlot()` or `reactive()`

```{r}
server <- function(input, output, session) {
    
    summary_table <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-03/athletes.csv') %>% 
        mutate(medal = factor(medal, levels = c("Bronze", "Silver", "Gold"))) %>% 
        group_by(medal, country, year) %>% 
        summarize(N = n())
    
    output$medals <- renderPlot(
        summary_table %>% 
            filter(country == input$country) %>% 
        ggplot() +
            aes(x = N, y = medal, fill = medal) +
            geom_col() +
            facet_wrap(~country) +
            theme_minimal() +
            scale_fill_manual(values = c("#AD7C42", "#D4D2C7", "#E4CC08"))
    )
}
```

- Above is an example of rendering a plot that is dependent on filtering by user input
- Notice how the data is read in in a separate step, in a non-reactive context -- so that this step doesn't have to be repeated every time the user changes the input!


### (3) shinyApp
```{r}
shinyApp(ui, server)
```

- Finally, just bring it all together into one `shinyApp()`!

## Reactivity 

- Reactive contexts allow R to be updated automatically whenever the user changes the input -- every time they drag the slider, as soon as they type one more letter, etc. 
  - renderText()
  - renderPlot()
  - reactive()

- Most reactivity happens in the server function, where input is turned into output 

- Input is read-only; you can't alter it within the server or anywhere else! Leave that to the user. 
  - It must be read in to a reactive context to allow for automatic updating

- Output must be rendered in a reactive context too! So the following code would throw an error:
```
output$greeting <- renderText("Hello, world!")
```

- Input and output that depend on each other are considered to be in a *reactive dependency* 

## Declarative programming 

- Shiny is based on *declarative* programming, so the mentality is a little different to R programming (which is *imperative*)

R: "paste together ‘hello’ and the user’s name, then send it to output$greeting" (IMPERATIVE)
Shiny: "if and when you need to send output$greeting to the browser, this is how you should do that" (DECLARATIVE)

R: "ERROR -- You misspelled your variable name as "greting" !!"
Shiny: "Okay, if I ever need to use "greting", now I know how to make it -- no problem here! Guess you just don't want any output!"

R: Order of lines matters! Top to bottom please.
Shiny: I use whatever line I need, based on reactive dependencies -- order doesn't really matter much to me.

R: Functions are useful when things get long, but just writing out what you want to happen in order usually does the trick.
Shiny: Functions help you understand and debug a lot and are basically essential!

# Data dashboard -- Eurovision

```{r}
library(tidyverse)
eurovision <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-17/eurovision.csv') %>% 
  filter(section == "grand-final")

head(eurovision)
```

(1) Figure out a filter statement that we can later let the user manipulate
```{r}
eurovision %>% 
  filter(year == 2021 & rank <= 3)
```

(2) Make a plot example to show the user

```{r}
eurovision %>% 
  filter(year == 2022 & rank < 4) %>% 
  ggplot(aes(x = reorder(artist_country, -total_points), y = total_points)) +
  geom_col()
```

(3) Add colors and labels
```{r}
eurovision %>% 
  filter(year == 2022 & rank < 4) %>% 
  ggplot(aes(x = reorder(artist_country, -total_points), y = total_points, fill = artist_country)) +
  geom_col(show.legend = F) +
  labs(
    y = "Total points",
    x = "Country"
  ) +
  theme_minimal()
```

Other options from RColorBrewer might be an option, but they don't have as many built-in levels as ggplot does! 
  
# (1) Add the graph 

Let's first just put this graph into a Shiny application, and have it be shown to the user. 

Make sure to: 
(A) Load in the data in a way that it won't have to be reloaded every time an input is changed
(B) Create the graph in a reactive context  
(C) Send the graph to output and show it to the user


## Spoilers

Hints: 
- Data should be read into the server function, but not in a reactive context (just in the body of the function)
- The graph should be created inside a reactive context like `renderPlot()`, inside the server function, and saved to an output object with a unique name 
- The graph should be displayed in the UI using `plotOutput()` and the name you have given the plot in the server


# (2) Collect user input

Now, let's let the user pick which year to display! 

```{r}
range(eurovision$year)
```

We can make this a `sliderInput()` or allow users to pick from a drop-down list with a `selectInput()`

Since there wasn't a contest in 2022, I'll use selectInput() so that I can remove this year from the options! 

After you add the ability to input, update the plot so that it reacts to the value given in the input.

## Spoilers

Make sure to:
(A) Add a `selectInput()` object to your UI, with numeric options from 2004:2022 excluding 2020 (you can type them out, or use a range for those years before 2019 and then add 2021 and 2022 to the vector)
(B) Change the filter command in your ggplot to be responsive to the input given, using the unique Id you gave the input 
(C) Use commas between parts of the UI ! 
 

# (3) Collect more user input

Let's also allow the user to select how many winners they want to see!

```{r}
eurovision %>% 
  drop_na(rank) %>% 
  group_by(year) %>% 
  summarize(max(rank))
```


Add another opportunity for user input, this time for the top X number of countries they'd like to see on the plot. This might be a nice `sliderInput()`

Then, update your plot to use this information as well!

Make sure to think about how many colors your palette has! You may have to set the max number lower unless you use a different palette. ggplot's built in palette will be fine though. 



## Spoilers

Be sure to: 
(A) Create a `sliderInput()` with how many countries should be displayed. 
(B) Update your filter command to include this input as well, using the ID you gave it. 
(C) You can use the number of countries to determine which ranks to show, i.e. those which ranked less than or equal to that number!

Also: Rotating axis labels will help when the user shows a lot of countries at once!
      theme(axis.text.x = element_text(angle = 45, vjust = 0.8))


# (4) Add a reactive title

Now, we could update the text on the screen to show what the user has inputted. 

Create a statement in `renderText()` that pastes together some text with the user's input, to describe the graph, e.g. 

"Eurovision Contest Placement in YEAR"
"Showing the top RANK countries"

Make sure to show this output to the user too!

## Spoilers 

Be sure to:
(A) Add a reactive context to the server using `renderText()` where you create an output of the text you want to show, using input values
(B) Show this output to the user in the UI using `textOutput()`

If you want, you can make two separate outputs, one as a title and one as a subtitle, so they easily show on separate lines. 


# Layout

Now we have a working Shiny app, let's craft the layout so that not everything is on top of each other!

First, we can try adding a title with a `titlePanel()` object. This goes in the UI and can include a regular string of text. 

Add this to the top of your ui -- maybe with a more creative title!
```
titlePanel("Eurovision Results"), 
```

Now, we can also add a sidebar panel and a main panel. 

For this, we need to first open a `sidebarLayout()` command in the UI, which contains two things: 
(1) a `sidebarPanel()` where our inputs will be 
(2) a `mainPanel()` for the plot! 

Take a look at the Shiny template for an example: 

```
library(shiny)
library(tidyverse)

# Define UI for application that draws a histogram
ui <- fluidPage(

    # Application title
    titlePanel("Old Faithful Geyser Data"),

    # Sidebar with a slider input for number of bins 
    sidebarLayout(
        sidebarPanel(
            sliderInput("bins",
                        "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30)
        ),

        # Show a plot of the generated distribution
        mainPanel(
           plotOutput("distPlot")
        )
    )
)

```

How could you add this to your Eurovision dashboard?

## Spoilers

(A) First, wrap all the code in your ui (other than the titlePanel)in `sidebarLayout()`
(B) Then, wrap all input parts in `sidebarPanel()` with a comma at the end!
(C) Finally, wrap the text that serves as a title for your plot, plus the plot itself in `mainPanel()`


# Conclusion

There are plenty of steps we could do now to beautify our Shiny app!

- find more color options or assign each country a set color, i.e. by making a hex code column
- adjust the width of our input section or the formatting of titles 
- add more content like a summary table or other type of graph

But we have a functioning Shiny app that allows the user to control both the year and the amount of countries they'd like to see. 

