---
title: "Guided Tidy Tuesday - Voice interactions in Star Trek"
author: "Julia Müller"
date: "24 8 2021"
output: html_document
---


# Exploring the data

Download from the TidyTuesday Github:
https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-17/readme.md

```{r}
library(tidyverse)
library(tidytext)

computer <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-17/computer.csv')
```

```{r}
head(computer, 20)

str(computer)

computer %>% 
  select(line, interaction)
```


## Figuring out what the column names mean
- What does "name" mean?
  ID - we can ignore that
- What's the difference between "line" and "interaction"?
  Line is the entire bit, including stage directions and sentences not said to computer
- ..."type" vs "pri_type"?
  Interaction type (can be multiple - text duplicated!), primary narrowed down to one
- ..."domain"? And some of the abbreviations, e.g. IoT
  Interaction domain - how is speech used? 
  IoT = Internet of Things - queries that activates or uses another piece of hardware
  Subdomains e.g. replicator (food & drink) or Turbolift
- ...the logicals: 
  nv_resp = nonverbal response
  is_fed = Federation computer
  error = Interaction resulted in error

Make sure to check out the codebook with much more information on what the categories mean here:
http://www.speechinteraction.org/TNG/TeaEarlGreyHotDatasetCodeBook.pdf


## Convert to factors
```{r}
str(computer)

summary(as_factor(computer$char))

computer <- computer %>% 
  mutate(across(c(char, type, pri_type, domain, sub_domain), as_factor))
```


# Analysis

## Who's the most talkative person?
```{r}
computer %>% 
  filter(char_type == "Person") %>% 
  count(char, sort = TRUE)
```

## How often is tea mentioned?
```{r}
computer %>% 
  filter(str_detect(line, "\\b[Tt]ea\\b")) # regular expression to find tea or Tea as a separate word (hence the would boundaries \\b) but not as part of another word, so not e.g. "steady"
```


# Graphs

## Speech types used by the computer vs the crew
Are human characters more likely than the computer character(s) to use speech that is classified as conversational?

### All speech types
Let's make a bar graph of speech types for the computer compared to the crew members:
```{r}
computer %>% 
  mutate(type = str_to_title(type)) %>% # to make sure all types start with a capital letter and remove spelling inconsistencies in the labeling
  group_by(char_type) %>% 
  count(type) %>% # create a summary table first
  mutate(type = fct_reorder(type, n)) %>% # reorder the factor levels so they're shown largest to smallest
  ungroup() %>% 
  ggplot() +
  aes(x = type, y = n, 
      fill = char_type) + # use different colours for computers and persons
  geom_col() + # because we calculated the summary earlier and the counts are present in the data, we use geom_col instead of geom_bar, which would sum up the relevant lines
  coord_flip() +
  facet_wrap(~ char_type, # create two plots, one for each char_type
             scales = "free") + # scale of both x- and y-axis of plots can be independent
  theme_minimal() +
  theme(legend.position = "none") # remove the legend
```

### Primary types only
Limiting this to the primary speech types, which means we need to remove duplicates with `distinct()`.
To sort the factor correctly, we need `reorder_within()` together with `scale_x_reordered()`:
```{r}
computer %>% 
  distinct(interaction, pri_type, # keep only unique interaction - primary type combinations to remove duplicates
           .keep_all = TRUE) %>% # keep all other columns - by default, only interaction and pri_type would be kept, but we need the other columns for the next steps
  group_by(char_type) %>% 
  count(pri_type) %>% # create a summary table first
  mutate(pri_type = reorder_within(pri_type, n, within = char_type)) %>% # reorder the factor levels so they're shown largest to smallest
  ungroup() %>% 
  ggplot() +
  aes(x = pri_type, y = n, 
      fill = char_type) + # use different colours for computers and persons
  geom_col() + # because we calculated the summary earlier and the counts are present in the data, we use geom_col instead of geom_bar, which would sum up the relevant lines
  coord_flip() +
  facet_wrap(~ char_type, # create two plots, one for each char_type
             scales = "free") + # scale of both x- and y-axis of plots can be independent
  theme_minimal() +
  scale_x_reordered() +
  theme(legend.position = "none") # remove the legend
```


## Speech types used by the crew members
```{r}
library(trekcolors) # colour palettes based on star trek colour schemes
```
More info:
https://github.com/leonawicz/trekcolors

Bar graph of speech types where the category labels and percentages are inside the bars:
```{r}
computer %>% 
  mutate(type = str_to_title(type)) %>%
  filter(char_type == "Person") %>% 
  count(type) %>% 
  mutate(type = fct_reorder(type, n),
         perc = round(n/sum(n) * 100, 2), # calculate the percentages, rounding to two positions after the comma
         perc_label = paste0(type, ": ", perc, "%"), # create the labels: speech type: percentage%
         place = if_else(perc >= 30, 1, 0)) %>% # where to place the labels: if the percentages is 30 or higher, the label should be inside the bar. Otherwise, outside.
  ggplot() +
  aes(x = type, y = perc, 
      fill = type) +
  geom_col() +
  geom_text(aes(label = perc_label,
                hjust = place)) + # hjust determines the placement of the text in the bar
  coord_flip() +
  theme_void() +
  scale_fill_trek("klingon", reverse = TRUE) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5), # both title and subtitle centered
        plot.subtitle = element_text(hjust = 0.5)) +
  labs(title = "Interaction types",
       subtitle = "Enterprise crewmembers") 
```


## Speech types used by the six most talkative crew members
Similar to the graph above, but with a `facet_wrap()` for a selection of crew members:
```{r}
computer %>% 
  filter(char_type == "Person") %>% 
  mutate(type = str_to_title(type),
         char = fct_lump_n(char, n = 6) # simplify the char factor: keep the six most frequent characters as they are but change everything else to "Other"
         ) %>% 
  filter(char != "Other") %>% 
  group_by(char) %>% 
  count(type) %>% 
  mutate(type = fct_reorder(type, n),
         perc = round(n/sum(n) * 100, 2),
         perc_label = paste0(type, ": ", perc, "%"),
         place = if_else(perc >= 20, 1, 0)) %>% 
  ggplot() +
  aes(x = type, y = perc, fill = type) +
  geom_col() +
  geom_text(aes(label = perc_label,
                hjust = place)) +
  coord_flip() +
  theme_void() +
  scale_fill_trek("klingon", reverse = TRUE) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  labs(title = "Interaction types",
       subtitle = "Enterprise crewmembers") +
  facet_wrap(~ char)
```


# Quick text analysis
Find and visualise the ten most characteristic words (by inverse document frequency) for each of the six most talkative characters: 
```{r}
computer %>% 
  filter(char_type == "Person") %>% 
  mutate(type = str_to_title(type),
         char = fct_lump_n(char, n = 6)) %>% 
  filter(char != "Other") %>% 
  unnest_tokens(output = "word", 
                input = interaction,
                token = "words") %>% # the data now has one word per row
  anti_join(stop_words) %>% # remove stopwords such as "the", "and", "of"...
  group_by(char) %>% 
  count(word) %>% 
  bind_tf_idf(word,
              char,
              n) %>% 
  slice_max(order_by = tf_idf, 
            n = 10, 
            with_ties = FALSE) %>% # limit to the ten words per character with the highest tf_idf
  mutate(word = fct_reorder(word, tf_idf)) %>% 
  ggplot() +
  aes(word, tf_idf, fill = char) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ char, scales = "free") +
  theme_minimal() +
  scale_fill_trek(palette = "lcars_alt") +
  theme(legend.position = "none")
```


# Links and resources

Cédric Scherer's tutorial on labels inside of bars:
https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/

Paper on this data: https://www.speechinteraction.org/TNG/AUTHORS_TeaEarlGreyHot_CHI2021.pdf

Our previous meetups on text analysis:
https://github.com/rladies/meetup-presentations_freiburg/tree/master/2020-07-Text_Analysis (explains tf-idf)
https://github.com/rladies/meetup-presentations_freiburg/tree/master/2020-09-Sentiment_Analysis
https://github.com/rladies/meetup-presentations_freiburg/tree/master/2021-06-22_TextAnalysis_ngrams

